Coalition Formation Algorithm of Prosumers in a Smart Grid Environment

### 2014-09-29 Article.tex
6 pages update

There seems to be a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}. Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management. The formation of coalitions inside the smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in some emergency situations (V2G) \cite{Ramchurn}. These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.

@@

\begin{center}
\begin{figure}
\includegraphics[scale=0.43]{figure2/Fig2}
\caption{Process diagram}
\label{Fig1}
\end{figure}
\end{center}

@@

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{l|p{4.5cm}}
  \hline 
  \\
  $ \mathcal{A}=\{a_{1},...,a_{N}\} $ & Set of agents \\
  $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
  $ \mathcal{T}_{i} = \{P_{i}(t_{0}),...,P_{i}(t_{K})\} $ & Time series of agent i available production values \\
  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of the production of the agent $i$ \\
  $ \phi \in [0,1] $ & Reliability threshold \\
  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value announced by coalition s \\
  $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
  $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
  $ \tau_{i}(s) \in \mathbb{R}$ & Decision rule for mapping overlapping into disjoint coalitions \\
  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
  $ CS $ & Set of all coalition structures \\
  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
  \\
  \hline
\end{tabular}

@@

\begin{figure}
 \centering
  \includegraphics[scale=0.9]{./figure9/util}
\caption{Utility function as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: a) In $ \{\phi \in [0,0.6], P^{MIN}\} $ when $ N_{COAL} $ is fixed. b) in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $. For readability, $ P^{MIN} $ is expressed in tenth of MW.}
\label{Fig2}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.8]{./figure5/coals}
  \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots). The diameters of the dots are proportional to the coalition sizes. The hatched zero utility zone corresponds to the \textit{"under-requirement space"}.}
  \label{Fig3}
\end{figure}

@@

\begin{figure}
 \centering
  \includegraphics[scale=0.35]{figure8/fig8.pdf}
  \caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $. Blue curves represent clique percolation, red curves, the random process, and the green ones, correlated coalitions.}
 \label{Fig4}
\end{figure}

>>>>

There seems to be a clear consensus on the benefit of having bidirectional communication flows between the prosumers and the grid, as demand side management relies on such an architecture \cite{WuHamedHuangBook2011}. Nevertheless as the number of active prosumers is expected to rise, it is safe to assume there is a need for more complex communication patterns (prosumers grouping together into so-called "coalitions") that will help to decrease the communication burden and satisfy the multiple requirements of the power grid management. Formation of coalitions inside a smart grid environment could be applied to various types of agents such as self sufficient micro-grids \cite{Pahwa}, sizable and adjustable virtual power plants (VPP) \cite{Braun, Ramchurn}, or fleet of electric vehicles that back up the grid in emergency situations (V2G) \cite{Ramchurn}. These are only a few use cases where the coalition of multiple agents can enhance the grid reliability and in the mean time reduce the communication burden.

@@

\begin{center}
\begin{figure}
\includegraphics[scale=0.37]{figure2/Fig2}
\caption{Process diagram}
\label{Fig1}
\end{figure}
\end{center}

@@

\begin{table}[t]
\centering
\scriptsize
\begin{tabular}{l|p{4.5cm}}
  \hline 
  \\
  $ P_{i}(t) \in \mathbb{R}\ \ (P_{s}(t)) $ & Instantaneous production of agent $i$ at time $t$ \\
  $ \mathcal{F}(P_{i})\ \ (\mathcal{F}(P_{S})) $ & Probability distribution of $i$'s production \\
  $ \phi \in [0,1] $ & Reliability threshold \\
  $ P^{MIN} \in \mathbb{R}^+ $ & Min value threshold \\
  $ P_{s}^{CRCT} \in \mathbb{R}^+ $ & Contract value announced by coalition s \\
  $ P_{\phi}(s) \in \mathbb{R}^+ $ & Max contract value under $ \phi $ constraint \\
  $ \mathcal{U}_{\phi,\ P^{MIN}}(s) $ & Utility function \\
  $ \tau_{i}(s) \in \mathbb{R}$ & Overlapping to disjoint coalitions mapping \\
  $ \epsilon \in [0,1] $ & Pruning threshold for the correlation graph \\
  $ N_{COAL} \in \mathbb{Z}^+ $ & Number of desired coalition \\
  $ CS $ & Set of all coalition structures \\
  $ \Theta(G) $ & Set of non overlaping cliques in graph G \\
  \\
  \hline
\end{tabular}

@@

\begin{figure}
 \centering
  \includegraphics[scale=0.8]{./figure9/util}
\caption{Utility as function of the parameters space $\{\phi \in [0,1], N_{COAL}\in \mathbb{Z}^+, P^{MIN} \in \mathbb{R}^+\}$: a) In $ \{\phi, P^{MIN}\} $ when $ N_{COAL} $ is fixed. b) in $ \{N_{COAL}, P^{MIN}\} $ for a fixed $ \phi $. For readability, $ P^{MIN} $ is expressed in tenth of MW.}
\label{Fig2}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.7]{./figure5/coals}
  \caption{Evolution of the coalitions for different values of $ P^{MIN} $ for clique percolation (blue dots) and random process (red dots). The diameters of the dots are proportional to the coalition sizes. The hatched zero utility zone corresponds to the \textit{"under-requirement space"}.}
  \label{Fig3}
\end{figure}

@@

\begin{figure}
 \centering
  \includegraphics[scale=0.3]{figure8/fig8.pdf}
  \caption{Evolution of the social welfare \textbf{a)} and the percentage of coalitions able to enter the market \textbf{b)} for different values of $ P_{MIN} $. Blue curves represent clique percolation, red curves, the random process, and the green ones, correlated coalitions.}
 \label{Fig4}
\end{figure}
